### 2016-02-22 math_modeling/class/notes/classnotes.tex
presentation for ethics

>>>>

\subsection{Arrow's Axioms}

Kenneth Arrow has several axioms that must be taken into account when modeling voting situations.
%page 185-187
\begin{enumerate}
\item The model should be transitive
\item The model should not consider irrelevant alternatives
\item The voters should be able to vote for whomever they choose
\item If every voter prefers one alternative to another, so does the whole society
\item There should be no case that leads to a dictatorship
\end{enumerate}

Among these axioms, (2) is the most logical of all. This axiom is stating that if we wish to know the
overall preference of $x$ vs $y$, then we \textbf{only} need to consider $x$ and $y$. In the general
case, this makes sense. The outcome should not be affected by candidate $z$, so we need not consider that
candidate.

\begin{quote}
\textit{\textbf{Example}} Consider the case where we have the following candidates:\\
\[\text{Rubio: A, Trump: B, Carson: C, Cruz: D, Kaish: E.}\]
\end{quote}
We can consider the following setup\\
\begin{center}
\begin{tabular}{c c c c c c}
A & B & C & D & E & E\\
D & E & B & C& B &C\\
E&D&E&E&D&D\\
C&C&D&B&C&B\\
B&A&A&A&A&A\\
\hline
18 & 12 & 10 & 9 & 4 & 2
\end{tabular}
\end{center}
\textbf{Plurality:} Under Plurality, A has the majority vote (18), so A wins.\\\\
\textbf{Single Runoff:} Under the single runoff model, the two candidates with the most 1st place votes
have a runoff. The winner of the runoff wins. The two top candidates here are A and B. In a single
runoff, A has 18 votes and B has 37, so B wins.\\\\
\textbf{Hare:} Under the Hare system, C wins.\\\\
\textbf{Bourda:} Under the Bourda method, D wins.\\\\
\textbf{Pairwise:} Under pairwise comparison, there are ${5 \choose 2} = 10$ comparisons.
In this case, E wins the election.
